Results section

### 2022-10-19 code/Result_section.Rmd
Updated file names

---
title: "Results section"
author: "Kaare"
date: "6/16/2022"
output: html_document
editor_options:
  chunk_output_type: console
---

@@

+-----------+-------------------------------------------------------------------------------+----------+----------+---------+
|           | PreExp 1                                                                      | PreExp 2 | PreExp 3 | NEC-FvT |
+===========+===============================================================================+==========+==========+=========+
| **Aim**   | To test feasibility of formula feeding with PICC line                         |          |          |         |
+-----------+-------------------------------------------------------------------------------+----------+----------+---------+
| Protocols | Testing a broad spectrum antibiotics mix in combination with formula feeding. |          |          |         |
|           |                                                                               |          |          |         |
|           | 4 groups                                                                      |          |          |         |
|           |                                                                               |          |          |         |
|           | 1.  CON-BF                                                                    |          |          |         |
|           | 2.  CON-FF                                                                    |          |          |         |
|           | 3.  AB-BF                                                                     |          |          |         |
|           | 4.  AB-FF                                                                     |          |          |         |
+-----------+-------------------------------------------------------------------------------+----------+----------+---------+
| N         | 38 (32 survived)                                                              |          |          |         |
+-----------+-------------------------------------------------------------------------------+----------+----------+---------+

@@

**Results**: `r inline_text(tbl1, {N})` mice were born to 6 different mothers. Of these 32 survived and were included in the analysis. Cytokines

```{r}
cytokines_1 <- read_excel(path = "data/processed/cytokines.xlsx") %>%  #Load in cytokine data for all experiments
  filter(exp_number == 1) %>%
  filter(tissue == "ileum") %>%
  mutate(
    group = recode_factor(  #Change the order of the groups to desired plot order
      group,
      "CON-BF"= "CON-BF",
      "CON-FF" = "CON-FF",
      "AB-BF"= "AB-BF",
      "AB-FF"= "AB-FF"
      ))

cytokines_1_long <- cytokines_1 %>%
  pivot_longer(cols = INFg:TNFa,
               names_to = "cytokines",
               values_to = "cytokine_concentration") %>%
  drop_na(cytokine_concentration)

normality_check <- cytokines_1_long %>%
  group_by(cytokines) %>%
  shapiro_test(cytokine_concentration) %>%
  select(cytokines, p) %>%
  filter(p < 0.05)

normality_plot <- cytokines_1_long %>%
  filter(cytokines %in% normality_check$cytokines ) %>%
  ggplot(aes(x = group,
             y = cytokine_concentration,
             color = group)) +
  geom_jitter() +
  facet_wrap(~ cytokines,
             scales = "free") +
  labs(title = "Non-normally distributed cytokines (Shapiro-Wilk Normality Test, p < 0.05)",
       y = "Concentration of cytokine in tissue (pg/mL)",
       x = "") +
  theme_bw()

normality_plot

normality_result_1 <- normality_check %>%
  knitr::kable(caption = "Non-normally distributed cytokines (Shapiro-Wilk Normality Test, p < 0.05)")
normality_result_1
```

@@

  tbl1_gt <- tbl1 %>%
  as_gt() %>%
  tab_header(title = md("**Proinflammatory cytokines**")) %>%
  opt_align_table_header("center") %>%
  tab_options(heading.title.font.size = 20,
              heading.background.color = "lightgrey",
              ) %>%
  tab_source_note(
    source_note = " * P-value indicates overall group differences"
  ) %>%
  tab_source_note(
    source_note = md("**Matrinem round 1**")
  ) %>%
  cols_width(label ~ px(100),
             stat_1 ~ px(150),
             stat_2 ~ px(150),
             stat_3 ~ px(150),
             stat_4 ~ px(150))
tbl1_gt
# Save and export files
# tbl1 %>%  #Save as excel file.
#   as_hux_table() %>%
#   huxtable::quick_xlsx(., file = "../Results/Round 1/ileum-cytokines.xlsx")
#
# tbl1 %>% #Save as image file.
# as_flex_table() %>%
#   flextable::save_as_image(path = "../../Results/Round 1/ileum_cytokines_table.png",webshot = "webshot")
```

@@

all_cytokines_patch
#ggsave('../../Results/Round 1/All Cytokines-boxplot.png',all_cytokines_patch)
```

@@

signif_cytokines_patch <- signif_cytokines + plot_annotation (
  title = 'Cytokine expression',
  subtitle = 'Statistically significant changes',
  tag_levels = 'A',
  caption = 'CON = Water, AB = Antibiotics, BF = Breast-Feeding, FF = Formula-Feeding. \n
  Pairwise comparisons with CON-BF using wilcoxon test',
  theme = theme(plot.title = element_text(hjust = 0.5, size=14, face="bold"),
                plot.subtitle = element_text(hjust = 0.5, face = "bold.italic"),
                plot.caption = element_text(hjust = 0.5, size=7,face = "bold")))
signif_cytokines_patch
#ggsave('../../Results/Round 1/Significant Cytokines-boxplot.png',signif_cytokines_patch)
```

```{r echo=FALSE, warning=FALSE}
#### Make regression model on long format
reg_mod_cyt_1 <- cytokines_1_long %>%
  group_by(cytokines) %>%
  do(lm(data = ., cytokine_concentration ~ maternal_treatment*type_of_feed) %>% tidy)  %>% filter(term!='(Intercept)') %>%
  ungroup() %>%
  filter(p.value < 0.05)

reg_mod_cyt_1 %>%
  select(-statistic) %>%
  knitr::kable(., caption = 'Significant effectors on cytokine expression', digits = 3)
```

Statistically significant group differences were observed for IL-6 (`r inline_text(tbl1, column = "p.value", variable = "IL_6")`), KCGRO (`r inline_text(tbl1, column = "p.value", variable = "KCGRO")`) and TNFa (`r inline_text(tbl1, column = "p.value", variable = "TNFa")`).

@@

-

    ## Will combining MAT with FF result in a more inflammatory intestine as measured by cytokine analysis in the ileum.

@@

# Experiment 2 - weight at birth.Summary: Offspring from antibiotics-treated mice were smaller at birth

```{r}
data <- read_excel("../Data/CleanData/weight_development.xlsx") %>%
  select(sample_id:cage, bodyweight_birthday_group_average, bodyweight_birthday_plus_24H_group_average, bodyweight_birthday_plus_48H_group_average, bodyweight_baseline, bodyweight_baseline_plus_24h, bodyweight_baseline_plus_48h) %>%
  rename(day_0 = "bodyweight_birthday_group_average",
         day_1 = "bodyweight_birthday_plus_24H_group_average",
         day_2 = "bodyweight_birthday_plus_48H_group_average",
         day_3 = bodyweight_baseline,
         day_4 = `bodyweight_baseline_plus_24h`,
         day_5 = `bodyweight_baseline_plus_48h`,) %>%
  pivot_longer(., cols = c(day_0:day_5),
                names_to = "time",
                values_to = "bodyweight") %>%
  mutate(received_antibiotics = recode_factor(received_antibiotics,
                                       yes = "MAT",
                                       no = "CON"))
  #%>%
  # #mutate(across(where(is_character),as_factor))

exp2.bw.density <- data %>%
  filter(exp_number == "2") %>%
  filter(time == 'day_0') %>%
  ggplot(aes(x = bodyweight, color = received_antibiotics)) +
    geom_density() +
  geom_vline(data = data %>%
              filter(exp_number == "2") %>%
  filter(time == 'day_0') %>%
  group_by(received_antibiotics) %>%
  summarise(grp.mean = mean(bodyweight)), aes(xintercept=grp.mean, color=received_antibiotics),
             linetype="dashed") +
  guides(fill = guide_legend(title = ""),
         colour = guide_legend(title = " ")) +
  labs(title = "", y = "n", x = "Bodyweight (g)") +
  theme_minimal()

exp2.bw.col <- data %>%
  filter(exp_number == "2") %>%
  filter(time == 'day_0') %>%
  ggplot(aes(x = received_antibiotics, y = bodyweight, fill = received_antibiotics)) +
  geom_bar(stat = "summary",
           fun = "mean") +
  labs(title = "", y = "Mean body weight", x = "") +
  theme_minimal() +  
  stat_compare_means(
    method = "t.test",
    label.y = 1.6,
    label.x.npc = 0.4) +
  theme(legend.position="none")

birth_weight_patch <- (exp2.bw.density + exp2.bw.col) + plot_annotation(
  title = 'Effect of Maternal Antibiotics Treatment on birth weight of offspring',
  subtitle = 'Experimental round 2',
  tag_levels = 'A',
  caption = 'MAT = Maternal Antibiotics Treatment, CON = Water Controls \n
  Statistics: non-paired two.sided t-test.'
)
birth_weight_patch
#ggsave("../Results/Weight/Figure_1_Birthweights.png", birth_weight_patch)

#add gt table with group results

```{r eval=FALSE, include=FALSE}
### Weight development during experiments   ###########

bw_development_feed <- data %>%
filter(exp_number %in% c("1","3","4")) %>%
filter(time %in% c("day_3", "day_5")) %>%
ggplot(aes(x = time ,
           y = bodyweight,
           fill = type_of_feed)) +
  facet_wrap(~ exp_number + type_of_feed,
             ncol = 2) +
  geom_bar(stat = "identity",
           position = "dodge") +
  guides(fill = guide_legend(title = " ")) +
  labs(
      caption = "Experimental round 2 not shown due to no formula feeding.",
    y = "Bodyweight (g)"
  ) +
  theme_bw()

bw_development_ab <- data %>%
filter(exp_number %in% c("1","2","3","4")) %>%
filter(time %in% c("day_3", "day_5")) %>%
ggplot(aes(x = time ,
           y = bodyweight,
           fill = received_antibiotics)) +
  facet_wrap(~ exp_number + received_antibiotics,
             ncol = 2) +
  geom_bar(stat = "identity",
           position = "dodge") +
  guides(fill = guide_legend(title = " ")) +
  labs(caption = "All mice recieved antibiotics in experimental round 4.",
    y = "Bodyweight (g)"
  ) +
  theme_bw()

bodyweight_develop_patch <- (bw_development_feed + bw_development_ab) + plot_annotation (
  title = 'Effect of Maternal Antibiotics Treatment on body weight development of offspring',
  subtitle = 'Faceted by experiment and Feeding or MAT-status',
  tag_levels = 'A',
  caption = 'Form = formular feed, Milk = breastfeeding.  
      MAT = Maternal Antibiotics Treatment, CON = Water Controls.')

bodyweight_develop_patch
#ggsave("../Results/Weight/Figure_2_bodyweight_development.png", bodyweight_develop_patch)

### Research Question 1: Body weight of offspring from water and antibiotic-treated mice at baseline

```{r style baseline bodyweight}
get_box_stats <- function(y, upper_limit = max(data$Bodyweight[data$Time == "Day2"]) * 1.0) {
  return(data.frame(
    y = 0.9 * upper_limit,
    label = paste(
      "n =", length(y), "\n",
      "Mean =", round(mean(y), 2), "\n" #,
      #"Median =", round(median(y), 2), "\n"
    )
  ))
}

my_comparisons <- list( c("Water", "AB"))

data %>%
  filter(Time == "Day2") %>%
  ggplot(aes(
    received_antibiotic, Bodyweight
  )) +
  geom_boxplot() +
  facet_wrap(~experiment_no,
             nrow = 1) +
  geom_dotplot(
     binaxis = "y", stackdir = "center", dotsize = 0.5) +
  stat_summary(
    fun.data = get_box_stats, geom = "text", hjust = 0.5, vjust = 0.9) +
  labs(
    title = "Body weight of offspring from antibiotic-treated vs water-treated mice",
    subtitle = "Results shown from each of the 4 experimental rounds",
    caption = "In Exp. 4 all mice received antibiotoics, hence no control.
    Comparison method: Student T-test",
    x = "Maternal Treatment",
    y = "Bodyweight at end of study (g)"
  ) +
  theme_bw() +
  stat_compare_means(
    method = "t.test",
    label.y = 3,
    comparisons = my_comparisons
    )
```

```{r weightgain plots}
############# Effect of Formula vs. Breastfeeding on Weight Gain" ###################

bodyweight_gain_BFvsFORM <- data %>%
  mutate(weightgain = weight_end - weight_baseline) %>%
  ggplot(aes(
    type_of_feed, weightgain
  )) +
  geom_boxplot() +
  facet_wrap(~experiment_no,
             nrow = 1) +
geom_dotplot(
     binaxis = "y", stackdir = "center", dotsize = 0.5) +
  stat_summary(
    fun.data = get_box_stats, geom = "text", hjust = 0.2, vjust = 0.9) +
  labs(
    title = "Effect of Formula vs. Breastfeeding on Weight Gain",
    subtitle = "Results shown from each of the 4 experimental rounds",
    caption = "In Exp. 2 all mice were breastfed.
    Comparison method: Student T-test",
    x = "Type of feed",
    y = "Body weight increase  (g)"
  ) +
  theme_bw() +
  stat_compare_means(
    method = "t.test",
    label.y = 3,
    comparisons = my_comparisons_2
    )
my_comparisons_2 <- list( c("Breastmilk", "Formula"))
bodyweight_gain_BFvsFORM
```

```{r weightgain plot 2}
############# Effect of  Antibiotocis on weightgain in Breastfed offspring" ################

bodyweight_gain_WATER_BF_vs_AB_BF <-
  data %>%
  mutate(weightgain = weight_end - weight_baseline) %>%
  filter(type_of_feed == "Breastmilk") %>%
  ggplot(aes(
    received_antibiotic, weightgain#, colour = mother
    )) +
  geom_boxplot()+
  geom_point()+
  facet_wrap(~experiment_no,
             nrow = 1) +
  labs(
    title = "Breastfed offspring +/- Maternal Antiobiotics Treatment"
  )
bodyweight_gain_WATER_BF_vs_AB_BF
```

#### Histological analysis

#### Cytokine expression

boxpltos

```{r}
boxplots_all_genes <- delta_gene_expr %>%
  ggplot(
    aes(x = groups,
        y = relative_fold_change,
        fill = groups)
  ) +
  geom_boxplot() +
  facet_wrap(~genes,
             ncol = 5,
             scales = "free") +
  labs(title = "Gene expression fold change",
       y = "Gene expression fold change",
       x = "Groups"
      )
boxplots_all_genes

#### Gut microbiome analysis

### Experiment 2 (E2)

#### Weight development

#### Histological analysis

#### Intestinal permeability

#### Cytokine expression

#### Gene expression

#### Gut microbiome analysis

### Experiment 4 (NEC-FvT)

The PreExps are now over and it is time to conduct the study. Having determined the optimal experimental protocol in PreExp 1-3 we are ready with the design.

#### Overview - Experimental Setup

#### Intestinal permeability - FITC

```{r FITC load data, include=FALSE}
fitc <- read_excel("../Data/CleanData/fitc.xlsx") %>%
  rename(values = `result_µg/ml_undiluted`) %>%
  drop_na(values)
```

Intestinal permeability is assessed with the FITC-Dextran assay. First I inspect the distribution of results.

```{r FITC check for normality}
fitc %>%
  filter(exp_number == 4) %>%
  group_by(group) %>%
  shapiro_test(values)
```

As results and highly skewed I try transforming the FITC results to log10.

```{r FITC try log transforming data}
fitc <- fitc %>%
  mutate(log_10_values = log10(values))  # Logtransform results

```{r FITC check for normality log transform}
fitc %>%
  filter(exp_number == 4) %>%
  group_by(group) %>%
  shapiro_test(log_10_values)
```

```{r FITC visual inspections of data distribution}
#Density plots non-log
density <- fitc %>%
  filter(exp_number == 4) %>%
  ggplot(aes(x = values, fill = group)) +
  geom_density() +
  facet_wrap(~ group,
             scales = "free")

#Jitter plots non-log
jitter <- fitc %>%
  filter(exp_number == 4) %>%
  ggplot(aes(x = group, y = values)) +
  geom_jitter() +
  facet_wrap(~ group,
             scales = "free")

#Density plots log
density <- fitc %>%
  filter(exp_number == 4) %>%
  ggplot(aes(x = , fill = group)) +
  geom_density() +
  facet_wrap(~ group,
             scales = "free")

#Jitter plots non-log
jitter <- fitc %>%
  filter(exp_number == 4) %>%
  ggplot(aes(x = group, y = values)) +
  geom_jitter() +
  facet_wrap(~ group,
             scales = "free")

#### Cytokine expression

#### Gene expression

#### Histological analysis

#### Gut microbiome analysis

>>>>

---
title: "Results section"
author: "Kaare"
date: "18/10/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

+-------------+-------------------------------------------------------------------------------+----------------------------------------------------------------+------------------------------------------------------------------------+-----------------------------------------------------------------+
|             | PreExp 1                                                                      | PreExp 2                                                       | PreExp 3                                                               | NEC-FvT                                                         |
+=============+===============================================================================+================================================================+========================================================================+=================================================================+
| **Aim**     | To test feasibility of formula feeding with PICC line                         | Selectively targeting gram positive vs. gram negative bacteria | Testing if there is an interaction-effect with feeding and antibiotics | Combining reults from PreExp 1-3 into final experimental setup. |
+-------------+-------------------------------------------------------------------------------+----------------------------------------------------------------+------------------------------------------------------------------------+-----------------------------------------------------------------+
| Protocols   | Testing a broad spectrum antibiotics mix in combination with formula feeding. | Groups                                                         | Groups                                                                 | Groups                                                          |
|             |                                                                               |                                                                |                                                                        |                                                                 |
|             | 4 groups                                                                      |                                                                |                                                                        |                                                                 |
|             |                                                                               |                                                                |                                                                        |                                                                 |
|             | 1.  CON-BF                                                                    |                                                                |                                                                        |                                                                 |
|             | 2.  CON-FF                                                                    |                                                                |                                                                        |                                                                 |
|             | 3.  AB-BF                                                                     |                                                                |                                                                        |                                                                 |
|             | 4.  AB-FF                                                                     |                                                                |                                                                        |                                                                 |
+-------------+-------------------------------------------------------------------------------+----------------------------------------------------------------+------------------------------------------------------------------------+-----------------------------------------------------------------+
| N           | 38 (32 survived)                                                              | 18                                                             | 32                                                                     | 26                                                              |
+-------------+-------------------------------------------------------------------------------+----------------------------------------------------------------+------------------------------------------------------------------------+-----------------------------------------------------------------+

@@

**Results**:

```{r include=FALSE}
preexp_1 <- read_excel("../data/processed/weight_development.xlsx") %>% 
  filter(exp_number == "1")

```{r}
#__________Load in cytokine data for all experiments________
cytokines_1 <- read_excel(path = "../data/processed/cytokines.xlsx") %>%  
  filter(exp_number == 1) %>% 
  filter(tissue == "ileum") %>% 
  mutate(
    group = recode_factor(  #Change the order of the groups to desired plot order
      group,
      "CON-BF"= "CON-BF",
      "CON-FF" = "CON-FF",
      "AB-BF"= "AB-BF",
      "AB-FF"= "AB-FF"
      ))
#___________________________________________________________

#__________ Pivot longer  ____________
cytokines_1_long <- cytokines_1 %>% 
  pivot_longer(cols = INFg:TNFa, 
               names_to = "cytokines", 
               values_to = "cytokine_concentration") %>% 
  drop_na(cytokine_concentration)
#___________________________________________________________
```

```{r}
#___________ Check cytokine distribution for normality, filter significant ones _____________________
normality_check <- cytokines_1_long %>% 
  group_by(cytokines) %>% 
  shapiro_test(cytokine_concentration) %>% 
  select(cytokines, p)
#___________________________________________________________

non_normal_cytokines <- normality_check %>% 
  filter(p < 0.05)

#___________ Jitterplot of all non-normally distributed cytokines _____________________
normality_plot <- cytokines_1_long %>%
  filter(cytokines %in% non_normal_cytokines$cytokines ) %>% 
  ggplot(aes(x = group, 
             y = cytokine_concentration,
             color = group)) + 
  geom_jitter() +
  facet_wrap(~ cytokines,
             scales = "free") + 
  labs(title = "Non-normally distributed cytokines (Shapiro-Wilk Normality Test, p < 0.05)",
       y = "Concentration of cytokine in tissue (pg/mL)",
       x = "") +
  theme_bw()
#___________________________________________________________

#___________ Jitterplot of all non-normally distributed cytokines _____________________
normality_result_1 <- normality_check %>% 
  knitr::kable(caption = "Normallity check of cytokine distribution (Shapiro-Wilk Normality Test, p < 0.05)")
#___________________________________________________________

names_of_cytokines <-unique(cytokines_1_long$cytokines)
```

`r nrow(preexp_1)` mice were born to 6 different mothers. Of these 32 survived and were included in the analysis.

`r nrow(cytokines_1)` samples were available for analysis of concentrations of the proinflammatory cytokines `r names_of_cytokines`. Concentrations were measured in 1-cm sections of proximal/distal ileum tissue using MSD Mesoscale Kits. Distributions were assessed for normality with Shapiro-Wilk Normality Test (see table below), and plot of data were made to assess distributions visually:

```{r echo=FALSE, fig.cap='Table:}
normality_result_1
```

```{r Normality plot, echo=FALSE, fig.cap='Non-normally distributed plots'}
normality_plot
```

@@

  tbl1_gt <- tbl1 %>% 
  as_gt() %>%
  tab_header(title = md("**Proinflammatory cytokines**")) %>%
  opt_align_table_header("center") %>% 
  tab_options(heading.title.font.size = 20,
              heading.background.color = "lightgrey",
              ) %>% 
  tab_source_note(
    source_note = " * P-value indicates overall group differences"
  ) %>%
  tab_source_note(
    source_note = md("**Matrinem round 1**")
  ) %>%
  cols_width(label ~ px(100),
             stat_1 ~ px(150),
             stat_2 ~ px(150),
             stat_3 ~ px(150),
             stat_4 ~ px(150))
# Save and export files
# tbl1 %>%  #Save as excel file.
#   as_hux_table() %>%
#   huxtable::quick_xlsx(., file = "../Results/Round 1/ileum-cytokines.xlsx")
# 
# tbl1 %>% #Save as image file.
# as_flex_table() %>%
#   flextable::save_as_image(path = "../../Results/Round 1/ileum_cytokines_table.png",webshot = "webshot")

The median concentration of the various cytokines are shown in the table below:

```{r, echo=FALSE}
tbl1_gt
```

@@

#all_cytokines_patch
#ggsave('../../Results/Round 1/All Cytokines-boxplot.png',all_cytokines_patch)
```

@@

signif_cytokines_patch <- signif_cytokines + plot_annotation (
  title = 'Cytokine expression',
  subtitle = 'Statistically significant changes',
  tag_levels = 'A',
  caption = 'CON = Water, AB = Antibiotics, BF = Breast-Feeding, FF = Formula-Feeding. \n
  Pairwise comparisons with CON-BF using wilcoxon test',
  theme = theme(plot.title = element_text(hjust = 0.5, size=14, face="bold"),
                plot.subtitle = element_text(hjust = 0.5, face = "bold.italic"),
                plot.caption = element_text(hjust = 0.5, size=7,face = "bold")))
#signif_cytokines_patch
#ggsave('../../Results/Round 1/Significant Cytokines-boxplot.png',signif_cytokines_patch)
```

The cytokines that were significantly different between groups are shown below:

```{r, echo= FALSE}
signif_cytokines_patch
```

A linear regression analysis calculates the factors responsible for the changes observed.

```{r echo=FALSE, warning=FALSE}
#### Make regression model on long format
reg_mod_cyt_1 <- cytokines_1_long %>% 
  group_by(cytokines) %>% 
  do(lm(data = ., cytokine_concentration ~ maternal_treatment*type_of_feed) %>% tidy)  %>% filter(term!='(Intercept)') %>% 
  ungroup() %>% 
  filter(p.value < 0.05)
```

```{r echo=FALSE, warning=FALSE}
reg_mod_cyt_1 %>% 
  select(-statistic) %>% 
  knitr::kable(., caption = 'Significant effectors on cytokine expression', digits = 3)
```

@@

@@

-   Will combining MAT with FF result in a more inflammatory intestine as measured by cytokine analysis in the ileum.
